---
title: "Zadanie1"
author: "Łukasz Berwid"
date: "1/6/2020"
output:
  html_document: default
  pdf_document: default
---


```{r libs, include=FALSE}
library("plyr")
library("dplyr")
library("ggplot2")
library("knitr")
library("plyr")
library("kableExtra")
#to render hraphs on Mac OS you need to install from github since my mac doesnt support x11
#devtools::intall_github("dcomtois/summarytools", ref = "no-x11-check")
library("summarytools")
```

```{r, include=FALSE}
st_css()   
knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(results = 'asis',     
                comment = NA, 
                prompt = FALSE, 
                cache = FALSE)

st_options(plain.ascii = FALSE,      
            style = "rmarkdown" ,       
            footnote = NA,            
            subtitle.emphasis = FALSE) 
```
## R Markdown

This data set contains weighted census data extracted from the 1994 and 1995 current population surveys conducted by the U.S. Census Bureau. The data contains demographic and employment related variables. [here](http://mlr.cs.umass.edu/ml/machine-learning-databases/census-income-mld/census-income.data.html).

Data URL [here](http://mlr.cs.umass.edu/ml/machine-learning-databases/census-income/census-income.data).

## 1. Download and load data

### Download data
```{r donwload}
 URL <- 'http://mlr.cs.umass.edu/ml/machine-learning-databases/census-income/census-income.data'
 CSVFilePath <- 'Zadanie1.csv'
 download.file(url=URL, destfile=CSVFilePath, method="libcurl")
```

### Load data
``` {r data_load, error=TRUE}
CSVFilePath <- 'Zadanie1.csv'
censusIncomeDataFrame <- read.csv(file=CSVFilePath, strip.white=TRUE)
colnames(censusIncomeDataFrame) <- c("age", "workclass", "fnlwgt", "education", "education-num", "marital-status", "occupation", "relationship", "race", "sex", "capital-gain", "capital-loss", "hours-per-week", "native-country", "class")

kable(summary(censusIncomeDataFrame))
```

## 2.
Columns that contains missing values
``` {r}
cols_with_missing_names <- colnames(censusIncomeDataFrame)[apply(censusIncomeDataFrame, MARGIN = 2, function(a) any(a=='?'))]
NameList <- cols_with_missing_names
idx <- match(NameList, names(censusIncomeDataFrame))
kable(colSums(censusIncomeDataFrame[,c(idx)]  == '?'), row.names = NA, col.names = 'missing count')
```

Total Number of missing values
``` {r}
kable(length(censusIncomeDataFrame[censusIncomeDataFrame=='?']), col.names = 'missing count')
```

## 3. Gender and age distribution

### Gender distribution chart
``` {r}
slices <- aggregate(censusIncomeDataFrame$sex,by=list(censusIncomeDataFrame$sex),FUN=length)
lbls <- c('Female', 'Male')
pct <- round(slices$x/sum(slices$x)*100)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie(slices$x, labels = lbls, main="Pie Chart of Countries")
```

### Age distribution historgram
``` {r}
hist(censusIncomeDataFrame$age, 
     main="Histogram for Age", 
     xlab="Age", 
     ylab="Count",
     xlim=c(0,100),
     breaks=10)
```

## 4. Table showing percetage count of native americans

``` {r}
natives_count <- length(which(censusIncomeDataFrame$`native-country` == "United-States")) + length(which(censusIncomeDataFrame$`native-country` == "OutlyingUS(Guam-USVI-etc)"))
natives_percentage <- (natives_count / count(censusIncomeDataFrame)) * 100

natives_percentage <- data.frame(total_count=count(censusIncomeDataFrame), natives_count=natives_count, natives_percentage=round(natives_percentage,2))

colnames(natives_percentage) <- rbind("total_count", "natives_count", "natives_percentage")
kable(natives_percentage)
```

## 5. Number of workhours per week by workclass

``` {r}
slices <- aggregate(censusIncomeDataFrame$`hours-per-week`,by=list(censusIncomeDataFrame$workclass),FUN=mean)
colnames(slices)<-rbind("workclass", "hours-per-week")

ggplot(slices, aes(x=slices$workclass, y=slices$`hours-per-week`)) + 
  geom_bar(stat = "identity") +
  xlab("Work class") +
  ylab("Mean number of hours per week") 
```


## 6. Distribution of average number of hours worked in private sector, for the group of people under age 30

``` {r}
under30_private <-subset(censusIncomeDataFrame, age < 30 & workclass == 'Private')

under30_private_summary <- under30_private %>% 
    group_by(`education`) %>% 
    summarise(count=n(), min = min(`hours-per-week`), max = max(`hours-per-week`), median = median(`hours-per-week`))

kable(under30_private_summary)

ggplot(under30_private_summary, aes(x = education, y = median, ymin = min, ymax = max)) + 
    geom_linerange() + 
    geom_pointrange()
```


## 7. Interesting plot

Relationship distribution by age group

``` {r}
censusIncomeDataFrame$age_group <- paste((round(censusIncomeDataFrame$age, -1)),  "-",  (round(censusIncomeDataFrame$age, -1) + 10))

suppressWarnings(ggplot(censusIncomeDataFrame, aes(relationship)) +
  ggtitle("Relationship status distribution per age group") +
  theme(plot.title = element_text(hjust = 0.5, face="bold")) +
  geom_histogram(stat="count") +
  xlab("relationship status") +
  ylab("Count") +
  facet_wrap("age_group"))
```

Data Summary

``` {r}
suppressWarnings(dfSummary(censusIncomeDataFrame, plain.ascii = FALSE, style = "grid", graph.magnif = 0.75, valid.col = FALSE))
```
